---
title: "Calculating and Comparing Bicycle Gears"
output: github_document
always_allow_html: true
---

**Credit:** The code and the considerable work in curating the `cassette_key.csv` dataset, in working out the logic underlying the functions in `bicycle_gears.R`, and the examples for each function was all done by David Roberts and described - along with commentary on why one front chainring (1x) is better than two (2x) in in his April 6 2016 blog post from [Cycling gear ratio calculator for R (or why you should just go buy a SRAM Eagle 1×12 group right now)](https://davidrroberts.wordpress.com/2016/04/06/cycling-gear-ratio-calculator-for-r-or-why-you-should-just-go-buy-a-sram-eagle-1x12-group-right-now/). 

I was interested in exploring different combinations, so naturally the first thing I did was search Google for "R bicycle gear ratios" and came across his blog post. Then decided to post on GitHub to make it easier for others to use.

Things you can do to contribute: 
- Improve & extend figures
- Develop a shiny app
- Finish converting into an R package

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```


* Functions are in `R/bicycle_gears.R`.
* Data is in `data/cassette_key.csv`.

## Read in data and functions


```{r}
library(dplyr)
cassette_key <- readr::read_csv('data/cassette_key.csv')

cassette_key ## todo make interactive table
```

```{r two-plus-two, exercise=TRUE}
source('R/bicycle_gears.R')
```

## Functions

### getCogs


```{r}

# SRAM Eagle
getCogs(brand = "SRAM", speeds = 12, small = 10, large = 50)

# Campy 13-speed road
getCogs("Campy", 13, 9, 42)
```

### Calculate Wheel Size

```{r wheelSize}
# For a typical 26" mountain bike setup with 2.1 tires
wheelSize(rimSize = 559, tireSize = 2.1, tireUnits = "inches")

# For a typical 700c road setup with 23c tires
wheelSize(622, 23, "mm")
```

### Generate tables of gear ratios, gear inches, speeds at different RPMs

```{r tables}
# Example for a typical road setup
# Standard 53-39 front
myRings <- c(39, 53)

# 11-speed 12-28 on the rear
myCogs <- getCogs("Shimano", 11, 12, 28)

# Running 700x23c
myWheel <- wheelSize(622, 23, "mm")

getGears(myRings, myCogs, myWheel)
```

## Make a plot comparing 1x and 2x options

```{r 1x_vs_2x}
twoby_rings <- c(39, 53)
# 11-speed 12-28 on the rear
oneby_rings <- 44
twoby_cogs <- getCogs("Shimano", speeds = 11, small = 12, large = 28)

oneby_cogs <- getCogs("Sram", speeds = 12, small = 10, large = 50)
# Running 700x23c
myWheel <- wheelSize(622, 23, "mm")

two <- cbind(type = '2x', bind_rows(getGears(twoby_rings, twoby_cogs, myWheel)))
one <- cbind(type = '1x', bind_rows(getGears(oneby_rings, oneby_cogs, myWheel)))
all <- rbind(one, two) %>% mutate(rings = as.factor(chainring), offset = ifelse(chainring == 53, 1, 0))


library(ggplot2)
# ggplot(data = all, aes(cog, gearInches, color = type, shape = rings)) + 
#   geom_point() +
#   geom_line()

ggplot(data = all, aes(type, gearInches, color = rings)) + 
  geom_point(aes(size = cog),alpha=0.5) +
  geom_line() +
  scale_color_grey() +
  geom_label(aes(label = cog), nudge_x = 0.1, size = 3, label.size= 0,alpha=0) +
  #scale_size_binned_area() +
  theme_bw() +
  coord_flip()

## todo: nudge 2x rings apart
## clean up fig
## make a shiny app
## make an R package
```

